Symbolic Parsing and\\ Automated Music Transcription

### 2021-05-03 main.tex
intro (debut)

% Parsing over infinite alphabet as optimal alignment computation 
% as edit-distance between string and language
% Symbolic Parsing
\title{Weighted Visibly Pushdown Automata\\ Application to Automated Music Transcription}
\author{Florent Jacquemard}
%\institute{INRIA \& CNAM, Paris, France\\
%\email{florent.jacquemard@inria.fr}}

@@

the problem
%quantitative parsing or symbolic parsing
of parsing of words over infinite input alphabet.

A weighted CFG associates a weight value to each of its derivation trees.
There exists some specialized parsing algorithms~\cite{nbest} 
to compute the $n$ best derivations of a given input word.

Such models and algorithm only deal with finite input alphabets.
Considering infinite alphabets can however be of practical interest, 
for instance when dealing with sequences of timestamped events.
...example of music transcription\dots

@@

We propose here models of weighted recognizers 
model of weighted CFG over infinite input alphabets %sets of terminal symbol
(but with finite sets of states and transitions rules). 
% non-terminals and production rules).
In the production rules, input symbols appear as variables (parameters)
and the weight associated to the transition rule is a function of these variables.
%
It is the approach of Symbolic Automata~\cite{VeanesdAntoniJACM}, 
except that the domain of weight values 
is not restricted to be Boolean, like  for the guards in the rules of SA, 
but can be an arbitrary commutative semiring (assuming some restrictions).

models: automata, transducer and pushdown automata with visibly restriction.
%wrt an edit distance
The latter corresponds in expressiveness
to a strict restricted model of weighted grammars. %VPA 
That permits to define a general approach for parsing 
over infinite input alphabets.

First one general edit-distance is defined by a weighted word 
transducer~\cite{Mohri}
%Symbolic automata transducers are extended models~\cite{VeanesdAntoniJACM}
%dealing with infinite set of input symbols...
value in a semiring...

@@

We follow the approach of~\cite{Mohri03EDWA} for the computation of distances
between words with weighted transducers, and propose models of weighted automata 
and weighted transducers over infinite alphabets.

>>>>

% Parsing over infinite alphabet as optimal alignment computation 
% as edit-distance between string and language
% 
\title{Symbolic Parsing and\\ Automated Music Transcription}
%\title{Weighted Visibly Pushdown Automata\\ Automated Music Transcription}
\author{Florent Jacquemard}
%\institute{INRIA \& CNAM, Paris, France\\
%\email{florent.jacquemard@inria.fr}}

@@

%the problem: quantitative parsing or symbolic parsing
%of parsing of words over infinite input alphabet.

Weighted parsing~\cite{Goodman99SemiringParsing,Nederhof03weightedParsing,MorbitzVogler19weighted-parsing} 
consists in analyzing a word in input \wrt a weighted language model, \eg a probabilistic context-free grammar (CFG).
In such a model, a weight value is associated to each transition rule, % production rule
and the rule's weights can be combined with a associative product operator $\otimes$ into 
the weight of a derivation (or \emph{abstract syntax tree}, \ie the result of an analyze).
A second operator $\oplus$, associative and commutative, 
is used to handle ambiguity of the model, 
by summing the weights of the possibly several (in general exponentially many) 
syntax trees associated to a given input word.
Intuitively, $\oplus$ aims at selecting, or ranking, the syntax trees.
The weight domain, equipped with these two operators is assumed, at minima, 
to form a \emph{semiring} $\Semiring$. %\cite{Eilenberg74automata}
%by ranking 
%making the weight domain a semiring.
Some efficient specialized parsing algorithms~\cite{Huang05kbest} have been proposed in this context 
in order to compute the $n$ best syntax trees of a given input word without having to enumerate them all.
Generally based on dynamic programming, these algorithms rely on 
additional algebraic properties of $\Semiring$.
-- see \eg~\cite{Huang05kbest} for some NLP applications.
%The extraction of $n$ best list is useful

** DEL ** 
%Parsing produces an analyze of an input word in the form of a parse tree 
%representing the derivation of a grammar.
In case of ambiguity, several analyses, in general an exponential number, 
are possible for the same input word. 
Weighted CFG associate a weight value to each of its derivation trees, 
based on weighted values associated to each production rules, 
and are therefore useful for the compact representation and the ranking of large sets of parse trees.
Moreover, there exist some efficient specialized parsing algorithms~\cite{Huang05kbest}
to compute the $n$ best derivations of a given input word.

The models and algorithms mentioned above deal with finite input alphabets.
Considering infinite alphabets can however be of practical interest, 
for instance when dealing with sequences of timestamped events.
(approach of music transcription by parsing~\cite{MCM19})

@@

We propose here some models of weighted recognizers 
%model of weighted CFG 
computing on words over infinite input alphabets %sets of terminal symbol
- but with finite sets of states and transitions rules. 
% non-terminals and production rules).
In their transition rules, input symbols appear as variables (parameters)
and the weight associated to the transition rule is a function of these variables.
%
This approach is close to the one of Symbolic Automata (SA)~\cite{VeanesdAntoniJACM}, 
except that the domain of weight values 
is not restricted to be Boolean, like for the guards in the rules of SA, 
but can be an arbitrary commutative semiring (assuming some restrictions).

The models defined here are finite automata, transducers and pushdown automata with visibly restriction.
%wrt an edit distance
The latter model of these automata computes on nested words~\cite{AlurMadhusudan09nested}, 
a structured format of words parenthesized with markup symbols, 
which can correspond to a linearization of trees.

The latter corresponds in expressiveness
to a strict restricted model of weighted grammars
and can be used in a general approach for parsing over infinite input alphabets.
\dots
First one general edit-distance is defined by a weighted word 
transducer~\cite{Mohri}
%Symbolic automata transducers are extended models~\cite{VeanesdAntoniJACM}
%dealing with infinite set of input symbols...
value in a semiring...

@@

We follow the approach of~\cite{Mohri03EDWA} for the computation of distances
between words with weighted transducers, and propose models of weighted automata 
and weighted transducers handling infinite alphabets.
